---
title: "What is the effect of hospital profit status on hospital prices?"
author: "Mo Singhal"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  pdf_document: 
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
fontsize: 12pt
geometry: margin = 1in
linestretch: 1.5
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes,
               scales, plotly, gganimate, cobalt, ivpack, stargazer, haven, ggthemes,
               gifski, magick, rdrobust, here, modelsummary, rddensity,
               MatchIt, cobalt, lfe, fixest, dotwhisker, kableExtra, knitr, Matching, )
```

```{r, include=FALSE}
load("470FinalProject.Rdata")

```

github repo: https://github.com/mosinghal1/470finalproject
\newpage

# Introduction
According to the American Hospital Association, about 3,000 hospitals, or fifty-eight percent of existing US hospitals in 2020, exist as non-profit entities (https://www.aha.org/infographics/2020-07-24-fast-facts-infographics, n.d.). In contrast, just over twelve hundred hospitals, or twenty-four percent of existing US hospitals in 2020, are ‘Investor-Owned For-Profit’ hospitals. Based on this finding, non-profit hospitals make up most hospitals in the United States and are double than for-profit hospitals. While this gap in number of for-profit and non-profit hospitals does exist, the distinction between for-profit and non-profit institutions in the US has been in contention in recent years.
Hospital care is distinctive in two important respects. First, the consumer is often not well informed about the quality of the service they are purchasing and frequently is less informed than the supplier, or the hospital. This is also known as adverse selection. Moreover, the consumer often cannot understand the quality of the good offered by the hospital. Second, consumers almost always pay out-of-pocket much less than the marginal cost of their care, and insurance companies or other parties foot the rest of the bill. Rather than provide care directly, governments primarily pay for care through insurance (Sloan, Picone, Taylor Jr, & Chou, 2001). In Dallas, for example, even after growing up there from the age of seven, I was unaware that the University of Texas Southwestern hospital was a non-profit hospital. I was also not aware of the Medical City of Dallas being a for-profit hospital owned by the hospital conglomerate HCA Healthcare. As I do not pay for my costs upfront, I, along with other healthcare consumers, do not deal with the true cost of care. 
With the many differences and ever-increasing similarities between for-profit and non-profit hospitals in the United States, this paper hopes to understand whether a hospital’s status affects the prices hospitals charge both insurance companies and consumers. Understanding the effect of profit status on hospital prices is incredibly relevant for the public good as well as public policy that communities, cities, counties, and states will use. If prices are the same between for-profit and non-profit hospitals, an argument could be made that non-profit hospitals are therefore not providing any added benefit to the community, thereby not deserving of their tax-exempt 501(c)(3) status. Moreover, if hospitals are not adding benefit to the community through decreased costs to consumers or prices charged for hospital procedures, then the hospitals could instead be taking away money from the community by not paying their fair share of property taxes, which they would have to pay if they were for-profit. With hospitals averaging approximately $192.8 million each year in net patient revenue, the taxes associated with these revenue amounts would be significant (Revenue Trends at U.S. Hospitals, n.d.). Per the information gathered, this paper hypothesizes there to be no effect on hospital prices due to a hospital’s profit status. 

# Literature Review
In the business sense, there are only four key differences between for-profit and non-profit institutions, which applies to hospitals and other companies operating in the healthcare space. The largest and most widely known difference is the tax status. For-profit hospitals are treated regularly, meaning that they must pay their fair share of taxes. Non-profit hospitals operate under a special designation from the Internal Revenue Service (IRS), 501(c)(3). As a 501(c)(3), the hospital is “exempt from paying federal income and unemployment taxes, and patrons who donate to them are allowed to claim a tax deduction for their contributions” (Kagan, 2022). Hospitals can then use this money to further support their purpose and support the public good. For-profit institutions and hospitals face another barrier on top of having to pay taxes as well, answering to shareholders and ever-increasing performance goals. Shareholders operate as part-owners of the institution and have a vested interest in making their money back. Moreover, upper management can be let go for not meeting performance goals set by ownership or the board of directors. In combination, shareholder pressure and increasing performance goals means that profits should always be trending upwards (Kagan, 2022). 
Serving as the third main difference between for-profits and non-profits, there does exist one up-side for-profit hospitals can take advantage of: access to capital and scalability. For-profit hospitals can use their ready access to capital to purchase more hospitals, allowing them to negotiate better prices with private insurance firms, decrease wasteful spending, and utilize economies of scale. Non-profits simply do not have the luxury of the access to capital, forcing them to move slower and lose out on a potentially cost-saving expansion (Cheney, 2017) . Lastly, non-profit hospitals must “serve the healthcare needs of the community” (Masterson, 2017). The 501(c)(3) designation is, in fact, based on this requirement that hospitals and other institutions must serve the community. Therefore, in addition to making enough money to survive and operate, non-profits must also continuously prove to regulatory authorities that they serve the community. Indeed, the question many consumers and communities ask is “whether the benefits to the community equal what taxpayers donate to these hospitals in the form of tax-exempt status” (Ofri, 2020). Many believe that the variation amongst hospitals in their benefit to the community does not warrant their tax exemptions. Notably, the city of Pittsburgh sued the University of Pittsburgh Medical Center (UPMC) in 2013 to remove UPMC’s tax-exempt status due to the city’s belief that the hospital was no longer operating in the best interest of the community and instead benefitting their upper management (Boren, 2013). In another case in 2015, a judge in New Jersey revoked the tax-exempt status for property taxes of a hospital system due to the hospital employing for-profit physician groups along with non-profit physicians. The judge went as far as to say “for purposes of the property tax exemption, modern nonprofit hospitals are essentially legal fictions” (Meikskins, 2015). 
While relevant, these four key differences have started to decrease in their effect on the operations of hospitals. Increasingly, non-profit hospitals have operated similarly to for-profit hospitals. In a report released in 2013 in an issue released by the journal Health Affairs, seven of the ten most profitable hospitals were non-profits. Only three, and not even the top three, of the ten most profitable hospitals were for-profit hospitals (Seven of top 10 most profitable U.S. hospitals are nonprofit, 2016).
Interestingly, much of the current existing literature acknowledges that hospital ownership affects the behavior of hospitals. A question many have asked is why hospitals, which majoritively act as financial institutions, would ever want to be non-profit hospitals in the first place. If non-profit hospitals are not maximizing profit, what are they maximizing instead? In an analysis done in 1988, Deneffe and Masson hypothesized that hospitals maximize utility with a focus on output while also considering profits. Deneffe and Masson, who analyzed Illinois hospitals between 1981 and 1983, found that cost-shifting allowed hospitals to earn money, but these non-profit hospitals reported and cared more about number of patients (Dennffe & Masson, 2002). In maximizing patients, or their ‘output,’ non-profit hospitals would be abiding by the purpose they must operate in for their tax-exempt status of benefiting the community. William J. Lynk, publishing in the Journal of Law and Economics, instead analyzed non-governmental prices of hospitals located in California throughout 1989. In comparing non-profit hospitals and for-profit hospitals, Lynk hypothesized, and later proved, that the low margins existing in non-profit hospitals at the time came from individual groups within hospitals taking more of the overall revenue, leading prices charged from the hospital to insurance companies and consumers to be extremely similar to both for-profit hospital prices and to the established profit-maximizing charge amounts (Lynk, 1995). 
A separate analysis by Yu-Chu et al published in The Journal of Health Care Organization, Provision, and Financing completed a meta-review of published papers. In the paper, a plurality of studies understanding the difference between ownership types of hospitals found that non-profits still focus on cost and revenue more than efficiency or community benefit, leading there to be truly trivial difference in the external or community effects of a hospital depending on the profit status (Yu-Chu, Eggleston, Lau, & H. , 2007). However, after standardizing the effects from each of the different papers, most papers indicated that for-profit hospitals have a higher cost. One paper studied in the meta-analysis by Yu-Chu et al. even found a 0.25 standard deviation towards higher costs at for-profit hospitals. Separately, the analysis also found that most papers saw little to no difference in revenue between for-profit and non-profit hospitals. In combination, with for-profit hospitals having higher costs but the same revenue as non-profit hospitals, one can assume that non-profit hospitals had more patients to make up for the lower costs. This finding is extremely perplexing as the same study found that non-profit hospitals still focus on maximizing profit, decreasing cost, and increasing revenue (Yu-Chu, Eggleston, Lau, & H. , 2007). This leads me to believe that while non-profit hospitals focus on the same outputs as for-profit hospitals, they still treat more patients. Whether on purpose or completely inadvertently, these non-profit hospitals do end up benefiting the community by seeing more patients than for-profit hospitals. 
An analysis of for-profit hospitals by Hyman et al. in 1997 found that for-profit hospitals, including those hospitals converting from non-profit to for-profit status ‘exploit’ their communities through “cream skimming, diversion, and exploitation”  (Hyman, 1997). While cream-skimming indicates hospitals locate in affluent areas for better-paying individuals and diversion refers to hospitals giving an adverse amount of money to their administrators and marketing, exploitation was defined as charging higher prices and inducing higher demand to make higher amounts of profit (Hyman, 1997). According to Sloan et al., there have been numerous instances of for-profit hospitals attempting to illegally skirt the system by charging more. Most notably, the Columbia-HCA, or the largest for-profit hospital firm, successfully billed the Medicare program in the Center for Medicare and Medicaid Services for services that were not provided and deemed inherently unnecessary (Sloan, Picone, Taylor Jr, & Chou, 2001). 
Interestingly, Sloan et al. also conducted an analysis of hospital market share, performance, and costs amongst non-profits and for-profit hospitals and found that for-profit hospitals had indeed increased prices and charges to Medicare. Moreover, for-profit hospitals do boast a higher cash-flow, allowing them to make more purchases of other hospitals and increase efficiency at a scale that non-profit hospitals cannot. This power, that non-profits have a “greater propensity to locate where they could make more money,” is also supported by Norton and Staiger et al in their analysis completed in 1994 (Norton & Staiger, 1994). 


# Data Description
The analyses performed in this report all utilize the Healthcare Cost Report Information System (HCRIS) made available by the Center for Medicaid and Medicare Services (CMS) website. After downloading the 1996 and 2010 datasets from the CMS website, many different variables were extracted and synthesized to form a final usable dataset. While the original HCRIS files provide an incredible amount of information, there are also certain problems that do exist, including the reporting of negative values. This could lead to errors in performed analyses, so the absolute value of these negative values was taken instead by the original author of the final dataset (McCarthy, n.d.). A second issue of hospitals having multiple duplicate report also exists in the data. While imperfect, an attempt was made to locate individual and unique reports for hospitals for each fiscal year. Again, the original author of the combined dataset made this attempt (McCarthy, n.d.). Finally, data available from the CMS website under the Provider of Services files was utilized to ascertain the profit status of hospitals based both on the national provider identifier (npi) number as well as the profit status by provider number. This available data was then merged with the final combined HCRIS data as new columns. 
After merging these data together, the dataset finally had the necessary information to test the hypothesis. Within the dataset, information such as the creation of each unique report (which ranged from 1998 to 2019) as well as the total charges, ancillary charges, and operating expenditure for each provider in each year existed. Additionally, the dataset included information such as Medicare payments, Medicare discharges, Medicaid discharges, total discounts, and provider identifying information. Of the hospitals available, nineteen percent are for-profit per the data available from the Provider of Services files. This matches up well with the data available form the American Hospital Association in which they indicated that twenty-four percent of all existing hospitals were ‘Investor-Owned For-Profit’ hospitals. While the American Hospital Association also included government-owned hospitals, the dataset used for the analyses in this report do not include data for community-owned hospitals. Instead, the remaining hospitals in this dataset are non-profit. 
To understand the data further, I made some figures showing individual correlations between variables in the data. In Figure 1, we see a line graph with two lines, indicating the hospital status. This figure specifically relates the total charges of hospitals from 1998 until 2018. As seen in Figure 1, the relative amounts of total charges between the for-profit hospitals and the non-profit hospitals stays relatively the same, with non-profit hospitals having a higher total charge amount for most of the time. This is an interesting aspect of the data as the figure coincides with the hypothesis that hospital prices are unaffected by a hospital’s profit status. However, it is also important to realize that total charges do not mean the same thing as total prices. While actual pricing data was not given by the CMS Hospital Cost Report (HCRIS) database, pricing data was calculated using the existing information. The step-by-step method for calculating the pricing data can be found in the empirical methods section. 
Figure 2 was also created based on the dataset showing a similar analysis on the total operating expenditure of the different hospital types across time. While there was no difference based on hospital profit status in the previous figure, Figure 2 shows an enormous difference between the for-profit hospitals and the non-profit hospitals. Specifically, the non-profit hospitals have nearly a double (or a $50,000,000) higher total operating expenditure than the for-profit hospitals on average each year. Over time, this gap only increased to a culmination of the data in 2018 showing that non-profit hospitals have nearly quadruple the amount of total operating expenditure compared to for-profit hospitals. At first glance, this result is unexpected. Operating expenditure would most likely include taxes and non-profit hospitals are exempt from paying federal and property tax, meaning that their total operating expenditure should be lower. Instead, we see that they are incredibly higher than the for-profit hospitals. While I do not have an adequate reason as to why we see this result, it could potentially be due to the size and number of large non-profit hospitals in the United States. Some of the largest hospitals in the US are non-profits including Mass General, Johns Hopkins Medical, the Mayo Clinic, and the Cleveland Clinic. With the enormous size of these hospitals, they would also have an incredibly large operating expenditure, leading to them being outliers that would increase the average total operating expenditures in each year. 
Lastly, in Figure 3, after calculating the price data, I was able to plot the yearly distributions of hospital prices based on the hospital profit status as well. In contrast to Figure 2 showing the large differentiation between for-profit and non-profit hospitals, Figure 3 shows almost no difference based on hospital profit status. Prices do increase as time goes on, but this could be based on inflation. While the data comes from an incredibly reputable source, one potential limitation is the lack of information regarding relative price and charge amounts in connection with inflation. Based on the data, I would say that the dollar amounts are not standardized to a certain year, meaning that the increase in prices over time could just be inflation and real dollar prices are staying the same across time. 


# Empirical Methods
In performing our analyses, the treatment variable was the profit status of the hospital. In the dataset, the profit status was indicated by a Boolean or dummy variable with for-profits having a ‘one value and non-profits having a ‘zero’ value. The output, or y-value, of most of the analyses was the price. Price was calculated by first calculating the discount factor, defined as one minus the ratio of total discounts to total charges for each report. Afterwards, a price numerator was found by summing the various individual charges including the in-patient charges, the ICU charges, and the ancillary charges and then multiplying by the discount factor, found above. To accurately understand the patient population, total discharges was subtracted by the Medicare discharges. Then, for each hospital cost report, the price numerator was divided by the patient population to find the per person price. After filtering for inadequate information and removing price outliers, a price variable was found in our data which would be our final output for most of the analyses. 
To test the prior hypothesis of whether a hospital’s profit status influences hospital prices, I chose to utilize the nearest neighbor matching analysis, specifically, the Mahalanobis nearest neighbor matching analysis. In doing so, M was set to 1, meaning that each control treatment data point found a singular nearest neighbor match in the non-control group. The goal for this analysis was to obtain the true causal inference of profit status on hospital prices, however, we do not have access to the counterfactual outcome, leading to an estimation required using the Mahalanobis nearest neighbor matching analysis. With the dearth of information available from both the CMS Hospital Cost Report (HCRIS) database as well as the CMS Provider of Services files providing hospital profits status, I believe that the selection on observables assumption will hold, meaning that there exists nothing unobserved which would affect the selection of hospital profit status and affects hospital prices for consumers and insurance companies. The second assumption that must be satisfied is common support, which assumes that someone of each type must be in both the treated and untreated groups. In the case of this analysis, a weighting analysis was accomplished and can be seen in Figure 4. The histograms of the untreated vs the treated group look what similar. This similarity indicates that the data passes the law of common support, allowing for the analyses to move forward and prove causal inference. 
After satisfying both necessary assumptions, a love plot (Figure 5) was created utilizing the Mahalanobis nearest matching as the input and understanding the effect of hospital profit status on four separate output variables: In Patient Charges, Total Medicare Payment, Price, and Total Operating Expenses of each hospital. Each of these outputs, other than price, which was explained earlier, came from the CMS Hospital Cost Report (HCRIS) database. 


# Results
Results come from both the Mahalanobis nearest neighbor matching analysis as well the love plot created by graphing the results of the nearest neighbor matching analysis. 
Based on the nearest neighbor matching analysis, the ATE is approximately -197.1 with a p-value of 2.12 x 10^-5. As we are computing the effect of hospital profit status on prices, a negative value would first indicate that as hospital profit status (a Boolean or dummy variable) increases, or becomes more for-profit, the hospital prices go down. Inversely, this would also mean that hospital prices increase as hospital profit status nears being a non-profit. With our outputs being hospital prices ranging from 0 to 60k dollars, an ATE of -197.1 would indicate that for every 1% increase of the input variable, or the profit status of the hospital, there would be a 197.1 dollar decrease in the hospital prices. This makes sense per our approximate range of prices, however, is at odds away from the hypothesis and the initial exploration of the data. As hospitals cannot be in-between hospital status (i.e., they cannot be half for-profit and half non-profit), the ATE would instead be a 100% increase from the ‘zero’ value of being a non-profit to the ‘one’ value of being a for-profit hospital. This would indicate that as the profit status increases by 100%, going from a non-profit to a for-profit hospital, there would be a 197.1 dollars times 100 or $19,710 decrease in prices. 
From our use of love plots in class and the love plot found in Figure 5 from the data in this analysis, we can also see the effects of hospital profit status on the four outcome variables: In-Patient Charges, Total Medicare Payments, Price, and Total Operating Expenses for each hospital report. Interestingly, the effect of the nearest neighbor matching is vast with the standardized mean difference shrinking incredible amounts after nearest neighbor matching. After the matching, the values for the standardized mean differences majoritively fall in the range of -0.1 to 0.1 standardized mean difference with only the total operating expenses falling just outside the -0.1 threshold. This small standardized mean difference, or SMD, indicates that the treatment effect is also relatively small. With the SMD being negative for the price effect, this coincides with the ATE of the Mahalanobis nearest neighbor matching as the ATE was negative as well. While we could not include 95% confidence intervals, the incredibly small p-value found in the nearest neighbor matching analysis leads us to believe that the results of the matching analysis are significant. 


# Conclusions
When starting this study, the goal was to understand the effects of hospital profit status on hospital prices: specifically, the prices charged to consumers and insurance companies. Through the completion of the analysis, there exists a strong likelihood that prices are affected by a hospital’s profit status and as shown by the nearest neighbor matching analysis, non-profits have higher prices. While this does go against the preliminary hypothesis, this is also generally surprising as for-profit hospitals are intended to maximize profits and would be expected to have the higher price amounts. Instead, we see this from the non-profits. In the data description section, we also found that total charges over time from 1998 until 2018 was similar between for-profit hospitals and non-profit hospitals, with non-profit hospitals having a slightly higher total charge amount for most of the available period. These findings, in combination, would give credence to the aforementioned point that the line between for-profit hospitals and non-profit hospitals is blurring. 
With for-profit hospitals charging lower prices than their non-profit counterparts, an argument could be made as to the value of the hospital to the community. As exemplified by the 2015 case in New Jersey against a hospital system that mixed for-profit and non-profit physician groups as well as the City of Pittsburgh suing the University of Pittsburgh Medical Center over their diversionary and extortionary practices, communities are starting to get wind of this lack of community value from tax-exempt non-profit hospitals as well. With the recouped tax that hospitals would (and should) be paying, communities can implement programs that would benefit the community, thereby doing the work that the non-profit hospitals should have been doing in the first place. When non-profit hospitals prioritize decreasing costs, increasing profits, and maximizing their own utility, they forsake their non-profit mandate of helping and supporting the community. 
Based on the results of the study, this paper recommends that individual municipalities and communities conduct their own analysis on the cost benefit analysis of each of their non-profit hospitals. In each case, a few questions should be asked: Does a hospital still deserve their 501(c)(3) tax-exempt status if they charge higher prices to consumers in the community? What ‘greater benefit’ to the community do they provide if charges and prices are still higher?

## Limitations
As mentioned earlier, one of the largest limitations is the lack of information regarding the real value of dollars throughout the twenty-year period for which data is available. 2018 dollars in no way can be exactly equal to the value of 1998 dollars, and therefore some standardization must occur to account for inflation. Unfortunately, there was no given information in how the costs, discounts, charges, and other financial information were represented, meaning that we do not know whether the financial information was standardized or not. While an intuitive analysis of the data suggests that the data was not standardized, an assumption as impactful as that would be outside of the scope of the purpose of this paper. 
Secondly, while the selection on observables assumption was met with the information provided in the CMS Hospital Cost Report (HCRIS) data, more data would have been useful to understand price changes. Most significantly, insurance prices and market share of hospitals would have painted a much clearer picture for the analyses completed in this paper to truly prove causal inference. Lastly, the Provider of Services data regarding hospital profit status had too few observations in comparison to the HCRIS data that already existed. For the analyses done in this paper, the profit status of each report was a vital piece of information, leading me to throw out valuable HCRIS data due to the lack of profit status. This adds a limitation on the study as some relevant hospital cost reports could have been used but were not. 

## Future Work
While this paper did prove that hospital prices were lower in for-profit hospitals, the relative size of the dataset included numerous non-profit hospitals for every one for-profit hospital. In the future, I would hope that a more evenly distributed dataset could be used to ensure that the results are statistically significant. Based off this work, I would also like to see municipalities and communities perform their own econometric analyses to understand the costs and benefits of non-profit hospital systems in their area of interest. Many times, as proven in this paper, the supposed benefit of a non-profit hospital, besides them majoritively being an incredibly large employer, also leads to less overall community benefit. In instances such as these, communities should follow the lead of the Judge in New Jersey and the University of Pittsburgh to challenge the tax-exempt status in the judiciary – and put the burden of proving their worth on the hospital systems rather than letting them lavish in their tax-free buildings free of charge. 
## Acknowledgements
This paper and the analyses that make up the report could not have been completed without the incredible foresight and educational prowess of Dr. Ian McCarthy. Much of the statistical, econometric, and data analytic information utilized and even taken for granted in this paper was taught by Dr. McCarthy to the author. In equal acknowledgement, the author would like to thank his Economics 470 SOS group of Grant Lawrence Britton, Rachel Songhua Ling, Senai Ghezai, and Michelle Pan, without whom this past semester would have been insurmountable and this report unfinished. 

# Works Cited

1.    (n.d.). Retrieved from https://www.aha.org/infographics/2020-07-24-fast-facts-infographics.
2.    Boren, J. (2013, March 20). Ravenstahl: Pittsburgh sues to remove UPMC's tax-exempt status.     Retrieved from TRIBLive: https://archive.triblive.com/news/ravenstahl-pittsburgh-sues-to-remove-upmcs-tax-exempt-status/
3.    Cheney, C. (2017, June 20). TOP 5 DIFFERENCES BETWEEN NFPS AND FOR-PROFIT HOSPITALS. Retrieved from HealthLeaders: https://www.healthleadersmedia.com/finance/top-5-differences-between-nfps-and-profit-hospitals
4.    Dennffe, D., & Masson, R. T. (2002). What do not-for-profit hospitals maximize? International Journal of Industrial Organization, 461-492.
5.    Hyman, D. A. (1997). Hospital conversions: fact, fantasy, and regulatory follies. J. Corp. L., 23, 741.
6.    Kagan, J. (2022, March 13). 501(c)(3) Organization. Retrieved from Investopedia: https://www.investopedia.com/terms/1/501c3-organizations.asp#:~:text=A%20501(c)(3)%20organization%20is%20a%20nonprofit%20organization,cruelty%20to%20children%20or%20animals.
7.    Lynk, W. J. (1995). Nonprofit Hospital Mergers and the Exercise of Market Power. The Journal of Law and Economics, 437-461.
8.    Masterson, L. (2017, May 25). Nonprofit, for-profit hospitals play different roles but see similar financial struggles. Retrieved from HealthcareDive: https://www.healthcaredive.com/news/nonprofit-for-profit-hospitals-play-different-roles-but-see-similar-financ/442425/
9.    McCarthy, I. M. (n.d.). Healthcare Cost Report Information System (HCRIS). Retrieved from GitHub: https://github.com/imccart/HCRIS
10.   Meikskins, R. (2015, July 7). Judge Terms Modern Nonprofit Hospitals a “Legal Fiction”. Retrieved from NonProfitQuarterly: https://nonprofitquarterly.org/judge-terms-modern-nonprofit-hospitals-a-legal-fiction/
11.   Norton, E., & Staiger, D. (1994). How hospital ownership affects access to care for the uninsured. The Rand journal of economics, 171-185.
12.   Ofri, D. (2020, February 20). Why are Nonprofit Hospitals So Highly Profitable? Retrieved from The New York Times: https://www.nytimes.com/2020/02/20/opinion/nonprofit-hospitals.html
13.   Revenue Trends at U.S. Hospitals. (n.d.). Retrieved from Definitive Healthcare: https://www.definitivehc.com/blog/revenue-trends-at-u.s.-hospitals#:~:text=According%20to%20data%20from%20the,at%20least%204%25%20each%20year.
14.   Seven of top 10 most profitable U.S. hospitals are nonprofit. (2016, May 2). Retrieved from Johns Hopkins Bloomberg School of Public Health: https://publichealth.jhu.edu/2016/nonprofit-hospitals-earn-substantial-profits
15.   Sloan, F. A., Picone, G. A., Taylor Jr, D. H., & Chou, S.-Y. (2001). Hospital ownership and cost and quality of care: is there a dime’s worth of difference? Journal of Health Economics, 1-21.
16.   Yu-Chu, S., Eggleston, K., Lau, J., & H. , C. (2007). Hospital Ownership and Financial Performance: What Explains the Different Findings in the Empirical Literature? The Journal of Health Care Organization, Provision, and Financing; Thousand Oaks, 41-68.



# Figures and Tables

\newpage
```{r Tot Charges, include=T, eval=T, fig.cap="Total Charges Based on Profit Status", fig.width=8, warning=F, echo=F}
profitvnonprofit_totcharges
```

\newpage
```{r Operating Expenditure, include=T, eval=T, fig.cap="Total Operating Expenditure Based on Profit Status", fig.width=8, warning=F, echo= F}
profitvnonprofit_operatingexp
```

\newpage
```{r Price Data by Year, include=T, eval=T, fig.cap="Hospital Prices per Year (W/ Major Outliers Removed)", fig.width=8, warning=F, echo=FALSE}
PriceDataYearGraph
```

\newpage

```{r, CommonSupportTest, include=T, eval=T, fig.cap="Histograms Summarizing Profit Status Boolean Variable", fig.width=8, warning=F, echo=F, message=F}
CommonSupportTest
```

\newpage

```{r Love Plot, include=T, eval=T, fig.cap="Love Plot of Output Variables Measuring Standardized Mean Differences", fig.width=8, echo=F}
FinalLovePlot
```

\newpage

```{r avgenrollment, include=T, eval=T, echo=F}
summary(m.nn.md)

```
Table 1: The Nearest Neighbor Matching Analysis


